---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
options(digits = 4, width = 160)
```

# covidestim
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

An R package for estimating population-level covid19 outcome probabilities (e.g.
hospitalization|infection, death|hospitalization, etc.), given the
age-distribution of a population of interest and severity estimates (by age,
sex, etc.) from various research groups.

Includes functions to extract country-level age distributions from the [UN World
Population Prospects
2019](https://population.un.org/wpp/Download/Standard/Population/) dataset, and
age-severity estimates from:

- [Salje et al. 2020](https://doi.org/10.1126/science.abc3517) (based on data
  from France)
- [Neher Lab](https://doi.org/10.1101/2020.05.05.20091363) (China)
- [Bi et al. 2020](https://doi.org/10.1101/2020.03.03.20028423) (Schenzhen,
  China)
- [Davies et al. 2020](https://doi.org/10.1101/2020.03.24.20043018) and [van
  Zandvoort et al. 2020](https://doi.org/10.1101/2020.04.27.20081711) (China,
  Diamond Princess)
- [O'Driscoll et al. 2020](https://doi.org/10.1101/2020.08.24.20180851) ensemble IFR (Infection Fatality Risk) estimates based on data from 45 countries.

## Installation

Install from GitHub with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("epicentre-msf/covidestim")
```

## Usage

#### Extract population age-distributions from UN World Population Prospects 2019

```{r}
library(covidestim)

covidestim::get_age_pop(iso = "AFG", format = "long") # AFG = Afghanistan
```

#### Derive population-wide covid19 outcome probability estimates

```{r}
# mean probability of death|infection in Afghanistan, based on age-severity
# estimates from different research groups
covidestim::get_p_Salje("AFG", p_type = "p_dead_inf")
covidestim::get_p_Neher("AFG", p_type = "p_dead_inf")
covidestim::get_p_LSHTM("AFG", p_type = "p_dead_inf")
```

#### Compare estimates of Pr(hospitalization|infection) from various research groups

```{r, warning=F, message=F, fig.width=9.5}
library(ggplot2)

p_hosp <- covidestim::compare_age_severity()

ggplot(p_hosp, aes(x = age_group, y = mean, color = group)) +
  geom_point(size = 3, position = position_dodge(width = 0.6)) +
  geom_linerange(aes(ymin = low95, ymax = upp95), position = position_dodge(width = 0.6)) +
  labs(x = "Age group", y = "Pr(hospitalization|infection)", col = "Group") +
  theme_bw()
```

#### Compare Infection Fatality Risk (IFR) of different countries

Get age-adjusted IFR estimates for countries according to ensemble
estimates by [O'Driscoll et al. 2020](https://doi.org/10.1101/2020.08.24.20180851).

```{r, warning=F, message=F, fig.width=10}
#get list of all countries with available population
pop <- get_pop_data()
cntrys <- pop$iso_a3
cntrys <- unique(cntrys)
cntrys <- cntrys[!is.na(cntrys)]

#compute IFRs, sort ascending, add continent
ifr <- compare_IFR(cntrys)
ifr$iso <- factor(ifr$iso, levels = ifr$iso[order(ifr$mn)])
ifr$continent <- countrycode::countrycode(ifr$iso, "iso3c", "continent")

#plot
ggplot(ifr[ifr$continent %in% c("Africa", "Asia"),]) +
  geom_point(aes(x = iso, y = 100*mn)) +
  geom_linerange(aes(x = iso, ymin = 100*low, ymax = 100*up)) +
  scale_y_log10(limits = c(0.05, 1)) +
  scale_x_discrete(guide = guide_axis(angle = 90, n.dodge = 2)) +
  labs(
    x = "Country",
    y = "IFR estimate (%)"
  ) +
  theme_bw() +
  theme(legend.position = "none") +
  facet_grid(cols = vars(continent), drop = T, scales = "free_x", space = "free_x")

ggplot(ifr[ifr$continent %in% c("Americas", "Europe", "Oceania"),]) +
  geom_point(aes(x = iso, y = 100*mn)) +
  geom_linerange(aes(x = iso, ymin = 100*low, ymax = 100*up)) +
  scale_y_log10(limits = c(0.05, 1)) +
  scale_x_discrete(guide = guide_axis(angle = 90, n.dodge = 2)) +
  labs(
    x = "Country",
    y = "IFR estimate (%)"
  ) +
  theme_bw() +
  theme(legend.position = "none") +
  facet_grid(cols = vars(continent), drop = T, scales = "free_x", space = "free_x")
```
